High-Dimensional Regularized Discriminant Analysis

### 2017-02-05 reports/paper-hdrda.Rnw
Shortened abstract/removed line numbers for Pattern Recognition.

\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib}
\usepackage{algorithm2e, url, lineno}

@@

\title{Computationally Efficient High-Dimensional Regularized Discriminant Analysis}

@@

Regularized discriminant analysis (\emph{RDA}), proposed by
\cite{Friedman:1989tm}, is a widely popular classifier that lacks
interpretability and is impractical for high-dimensional data sets.  Here, we
present an interpretable and computationally efficient classifier called
high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample,
high-dimensional setting. For \emph{HDRDA}, we show that each training
observation, regardless of class, contributes to the class covariance matrix,
resulting in an interpretable estimator that \emph{borrows} from the pooled
sample covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to
a classifier in a reduced-feature space with dimension approximately equal to
the training sample size. As a result, the matrix operations employed by
\emph{HDRDA} are computationally linear in the number of features, making the
classifier well-suited for high-dimensional classification in practice. We
demonstrate that \emph{HDRDA} is often superior to several sparse and
regularized classifiers in terms of classification accuracy with three
artificial and six real high-dimensional data sets. Also, timing comparisons
between our \emph{HDRDA} implementation in the {\tt sparsediscrim} R package
and the standard \emph{RDA} formulation in the {\tt klaR} R package demonstrate
that as the number of features increases, the computational runtime of
\emph{HDRDA} is drastically smaller than that of \emph{RDA}.

@@

\linenumbers

@@

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times
  1}$ and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times
  p}^{>}$, $k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are
typically unknown, we assign an unlabeled observation $\bm x$ to class
$\omega_k$ with the sample \emph{QDA} classifier
\begin{linenomath*}
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
\end{align}
\end{linenomath*}
where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum-likelihood
estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled
sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{
  \bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{linenomath*}
\begin{align}
	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
\end{align}
\end{linenomath*}
is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
\emph{LDA} classifier. We omit the log-determinant because it is constant across
the $K$ classes.

@@

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and provides reduced variability for
$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix
regularization applies a shrinkage factor $\gamma > 0$, so that
\begin{linenomath*}
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
\end{linenomath*}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{linenomath*}
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) \coloneqq (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
\end{linenomath*}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{linenomath*}
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
\end{linenomath*}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, we see that $\lambda$ weights the
contribution of each of the $N$ observations in estimating $\bm \Sigma_k$ from
all $K$ classes rather than using only the $n_k$ observations from a single
class. As a result, we can interpret \eqref{eq:sig-lambda-alternative2} as a
covariance-matrix estimator that borrows from $\widehat{\bm \Sigma}$ in
\eqref{eq:pooled-cov} to estimate $ \bm \Sigma_k$.

@@

As we have discussed above, several eigenvalue adjustment methods have been
proposed that increase eigenvalues (approximately) equal to 0. To further
improve the estimation of $\bm \Sigma_k$ and to stabilize the estimator's
inverse, we define the eigenvalue adjustment of
\eqref{eq:sig-lambda-alternative} as
\begin{linenomath*}
\begin{align}
	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
\end{linenomath*}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
\emph{HDRDA} classifier as
\begin{linenomath*}
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
\end{linenomath*}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
{\tt sparsediscrim} R package, which is available on CRAN.

@@

When $\alpha_k = 1 - \gamma$, \eqref{eq:hdrda-cov} resembles the biased covariance-matrix estimator
\begin{linenomath*}
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k^{(RDA)}(\lambda) + \gamma \frac{\tr\left\{\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)\right\}}{p} \bm I_p \label{eq:sig-rda}
\end{align}
\end{linenomath*}
employed in the \emph{RDA} classifier, where $\widehat{\bm
  \Sigma}_k^{(RDA)}(\lambda)$ is a pooled estimator of $\bm \Sigma_k$ and
$\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of
\eqref{eq:sig-rda} towards $\bm I_p$ weighted by the average of the eigenvalues
of $\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)$. Despite the similarity of
\eqref{eq:sig-rda} to the \emph{HDRDA} covariance-matrix estimator in
\eqref{eq:hdrda-cov}, the \emph{RDA} classifier is impractical for
high-dimensional data because the inverse and determinant of \eqref{eq:sig-rda}
must be calculated when substituted into \eqref{eq:qda}. Furthermore,
\eqref{eq:sig-rda} has no clear interpretation.

@@

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the eigendecomposition of
$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
q \le p$. Then, we have
\begin{linenomath*}
\begin{align}
	\tilde{\bm \Sigma}_k &= \bm U
    \begin{bmatrix}
      \bm W_k & \bm 0 \\
      \bm 0 & \gamma \bm I_{p-q}
    \end{bmatrix}
    \bm U^{T}, \quad k = 1, \ldots, K,\label{eq:rda-matrix}
\intertext{where}
\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
\end{align}
\end{linenomath*}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
= \bm 0_{p \times (p - q)}$. Hence,
\begin{align*}
  \bm U^{T} \widehat{\bm \Sigma}_k \bm U = \begin{bmatrix}
    \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 & \bm 0 \\
    \bm 0 & \bm 0_{p-q}
  \end{bmatrix}, \quad k = 1, \ldots, K.
\end{align*}
Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U^{T}
\widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
\eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
\end{proof}

@@

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the decision rule in \eqref{eq:hdrda} is equivalent to
\begin{linenomath*}
\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
\end{align}
\end{linenomath*}
\end{thm}
\begin{proof}
From \eqref{eq:rda-matrix}, we have that
\begin{align*}
  \tilde{\bm \Sigma}_k^{+} = \bm U \begin{bmatrix}
    \bm W_k^{-1} & \bm 0 \\
    \bm 0 & \gamma^{+} \bm I_{p-q}
  \end{bmatrix}
  \bm U^{T}
\end{align*}
and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$, $k = 1, \ldots,
K$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
\begin{linenomath*}
\begin{align*}
	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
  &+ (p - q) \log \gamma.
\end{align*}
\end{linenomath*}
Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
\end{proof}

@@

\begin{proposition}\label{proposition:hdrda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
and
\begin{linenomath*}
\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - n_k^{-1} \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
    \intertext{where}
    \bm Q_k &= \bm I_{n_k} + n_k^{-1} \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
    \intertext{and}
    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
\end{align}
\end{linenomath*}
\end{proposition}
\begin{proof}
First, we write $\bm W_k = n_k^{-1} \alpha_k (1 - \lambda) \bm U_1^{T} \bm
X_k^{T} \bm X_k \bm U_1 + \bm \Gamma_k$. To calculate $|\bm W_k|$, we apply
Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$, where
$\bm A \in \mathbb{R}_{a \times a}^{>}$, $\bm B \in \mathbb{R}_{a \times b}$,
$\bm T \in \mathbb{R}_{b \times b}^{>}$, and $\bm C \in \mathbb{R}_{b \times
  a}$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm
U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we
have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv}
follows from the well-known Sherman-Woodbury formula \citep[Theorem
  18.2.8]{Harville:2008wja} because $(\bm A + \bm B \bm T \bm C)^{-1} =
\bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
\bm A^{-1}$.
\end{proof}

>>>>

\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib}
\usepackage{algorithm2e, url}

@@

\title{High-Dimensional Regularized Discriminant Analysis}

@@

Regularized discriminant analysis (\emph{RDA}) is a widely popular classifier that lacks interpretability and is impractical for high-dimensional data. Here, we present a competing classifier called high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample, high-dimensional setting. For \emph{HDRDA}, we show that each training observation, regardless of class, contributes to the class covariance matrix, resulting in an interpretable estimator that \emph{borrows} from the pooled sample covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to a classifier in a reduced-feature space with dimension approximately equal to the sample size. Consequently, the matrix operations employed by \emph{HDRDA} are linear in the number of features. Timing comparisons between \emph{HDRDA} and RDA establish that as the number of features increases, the computational runtime of \emph{HDRDA} is drastically smaller than that of \emph{RDA}. Also, we demonstrate that \emph{HDRDA} often attains superior classification accuracy to several sparse and regularized classifiers using three artificial and six real high-dimensional datasets.

@@

@@

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times
  1}$ and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times
  p}^{>}$, $k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are
typically unknown, we assign an unlabeled observation $\bm x$ to class
$\omega_k$ with the sample \emph{QDA} classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum-likelihood
estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled
sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{
  \bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
\emph{LDA} classifier. We omit the log-determinant because it is constant across
the $K$ classes.

@@

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and provides reduced variability for
$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix
regularization applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) \coloneqq (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, we see that $\lambda$ weights the
contribution of each of the $N$ observations in estimating $\bm \Sigma_k$ from
all $K$ classes rather than using only the $n_k$ observations from a single
class. As a result, we can interpret \eqref{eq:sig-lambda-alternative2} as a
covariance-matrix estimator that borrows from $\widehat{\bm \Sigma}$ in
\eqref{eq:pooled-cov} to estimate $ \bm \Sigma_k$.

@@

As we have discussed above, several eigenvalue adjustment methods have been
proposed that increase eigenvalues (approximately) equal to 0. To further
improve the estimation of $\bm \Sigma_k$ and to stabilize the estimator's
inverse, we define the eigenvalue adjustment of
\eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
\emph{HDRDA} classifier as
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
{\tt sparsediscrim} R package, which is available on CRAN.

@@

When $\alpha_k = 1 - \gamma$, \eqref{eq:hdrda-cov} resembles the biased covariance-matrix estimator
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k^{(RDA)}(\lambda) + \gamma \frac{\tr\left\{\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)\right\}}{p} \bm I_p \label{eq:sig-rda}
\end{align}
employed in the \emph{RDA} classifier, where $\widehat{\bm
  \Sigma}_k^{(RDA)}(\lambda)$ is a pooled estimator of $\bm \Sigma_k$ and
$\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of
\eqref{eq:sig-rda} towards $\bm I_p$ weighted by the average of the eigenvalues
of $\widehat{\bm \Sigma}_k^{(RDA)}(\lambda)$. Despite the similarity of
\eqref{eq:sig-rda} to the \emph{HDRDA} covariance-matrix estimator in
\eqref{eq:hdrda-cov}, the \emph{RDA} classifier is impractical for
high-dimensional data because the inverse and determinant of \eqref{eq:sig-rda}
must be calculated when substituted into \eqref{eq:qda}. Furthermore,
\eqref{eq:sig-rda} has no clear interpretation.

@@

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the eigendecomposition of
$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
q \le p$. Then, we have
\begin{align}
	\tilde{\bm \Sigma}_k &= \bm U
    \begin{bmatrix}
      \bm W_k & \bm 0 \\
      \bm 0 & \gamma \bm I_{p-q}
    \end{bmatrix}
    \bm U^{T}, \quad k = 1, \ldots, K,\label{eq:rda-matrix}
\intertext{where}
\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
\end{align}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
= \bm 0_{p \times (p - q)}$. Hence,
\begin{align*}
  \bm U^{T} \widehat{\bm \Sigma}_k \bm U = \begin{bmatrix}
    \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 & \bm 0 \\
    \bm 0 & \bm 0_{p-q}
  \end{bmatrix}, \quad k = 1, \ldots, K.
\end{align*}
Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm U^{T}
\widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
\eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
\end{proof}

@@

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the decision rule in \eqref{eq:hdrda} is equivalent to
\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
\end{align}
\end{thm}
\begin{proof}
From \eqref{eq:rda-matrix}, we have that
\begin{align*}
  \tilde{\bm \Sigma}_k^{+} = \bm U \begin{bmatrix}
    \bm W_k^{-1} & \bm 0 \\
    \bm 0 & \gamma^{+} \bm I_{p-q}
  \end{bmatrix}
  \bm U^{T}
\end{align*}
and $|\tilde{\bm \Sigma}_k| = \gamma^{p-q}| \bm W_k |$, $k = 1, \ldots,
K$. Therefore, for all $\bm x \in \mathbb{R}_{p \times 1}$, we have that
\begin{align*}
	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
  &+ (p - q) \log \gamma.
\end{align*}
Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
\end{proof}

@@

\begin{proposition}\label{proposition:hdrda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
and
\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - n_k^{-1} \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
    \intertext{where}
    \bm Q_k &= \bm I_{n_k} + n_k^{-1} \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
    \intertext{and}
    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
\end{align}
\end{proposition}
\begin{proof}
First, we write $\bm W_k = n_k^{-1} \alpha_k (1 - \lambda) \bm U_1^{T} \bm
X_k^{T} \bm X_k \bm U_1 + \bm \Gamma_k$. To calculate $|\bm W_k|$, we apply
Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$, where
$\bm A \in \mathbb{R}_{a \times a}^{>}$, $\bm B \in \mathbb{R}_{a \times b}$,
$\bm T \in \mathbb{R}_{b \times b}^{>}$, and $\bm C \in \mathbb{R}_{b \times
  a}$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm
U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we
have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv}
follows from the well-known Sherman-Woodbury formula \citep[Theorem
  18.2.8]{Harville:2008wja} because $(\bm A + \bm B \bm T \bm C)^{-1} =
\bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
\bm A^{-1}$.
\end{proof}
